### 2013-09-23 intro.tex
feedback from a real expert....

%\includegraphics[width=\linewidth]{python_soul}%
%\flushright%
%Randall Munroe, \emph{xkcd.com}
\end{minipage}%
\\[2cm]
In computer vision research, the goal is to automatically extract information
from a given image of image sequence.  In particular discerning semantic
information, that is interpreting an image, is a prominent research topic.
While much progress has been made in recent years, computer vision systems still lag behind
human vision in most tasks that require semantic information. These tasks can often be formulated
in terms of \emph{semantic classes}, meaning categories of parts, objects or scenes.
Examples include answering questions such as ``Is this a picture of a beach?'', ``How many
cars are there in this image?'' or even ``What objects lie on the table?''.
These questions illustrate a range of possible tasks involving semantic categories,
such as classifying images of single objects, localization and counting of object classes,
and parsing a scene fully into objects and object classes together with their relations.
%
While humans can distinguish tens of thousands of object classes, and have
little trouble in interpreting complex scenes, current methods are often
restricted to a much smaller number of classes and only have limited
capabilities to model interactions or relations.  We believe that
\emph{context} is one of the most important cues when it comes to classifying
objects, and therefore understanding scenes. Consequently, we target dense
labeling of scenes, taking object relations into account.  The task of densely
labeling image pixels into classes is called object class segmentation or
semantic segmentation.
\pagebreak

We choose this task in particular for several reasons:
\begin{itemize*}
    \item Pixel-level annotation provides highly detailed information about the scene.
    \item Joint estimation of multiple classes allows for the use of context.
    \item In contrast to category-agnostic segmentation approaches, object class segmentation
        has an unambiguous true labeling.
    \item A variety of manually annotated datasets is publicly available.
\end{itemize*}
%
Applications of semantic segmentation and scene understanding include automatic
image interpretation for retrieval, autonomous driving and mobile robotics.
%While these are already important applications, we expect that with the
%abundance of camera data, and the proliferation of mobile computing, even
%more applications will emerge soon.
Besides these applications, due to the abundance of camera data and the
proliferation of mobile computing, we expect semantic annotation of images to
be a key component in future technologies.

@@

One of the bottlenecks in learning object class segmentation and semantic
segmentation is the availability of training data.  While unlabeled image data,
and even data with semantic
``tags'' is available in practically unlimited quantities, semantic annotation
on pixel level is only available through laborious manual annotation. 
Chapter~\ref{ch:semi_supervised} introduces an approach to cope with this
shortcoming of object class segmentation approaches by introducing a method to
find segmentations automatically from image level annotations.

The main part of this thesis investigates the use of structured learning~\citep{taskar2003max, tsochantaridis2006large}
algorithms to the task of semantic image segmentation. Both topics have received
much attention in the computer vision and machine learning communities
lately~\citep{ladicky2009associative, krahenbuhl2012efficient,
branson2013efficient, blake2011markov}.
Unfortunately, learning structured prediction in computer vision applications
is still little understood.
We focus on the use of conditional random fields~(CRFs), which have shown
promising  results for computer vision applications. Using the paradigm of structural
support vector machines, it is possible to learn conditional random fields to
directly minimize a loss of interest. In particular, CRFs
allow to combine different cues, possibly produced by other approaches, in a
principled manner. One of the main difficulties with CRF approaches to computer
vision problems is that context in images is usually represented as a
neighborhood graph.  These graphs, by nature, contain many cycles, making
inference intractable in general. Consequently, learning algorithms have to
rely on approximate inference, with often unknown consequences to learning.

@@

\section{List of Contributions}
This thesis contains the following contributions:
\begin{itemize}
\item Introducing a clustering algorithm that improves upon widely used
    approaches from the literature. Our algorithm yields better clusterings in
    terms of known classes on a wide range of standard datasets.
\item Demonstrating a scalable algorithm for weakly supervised object class
    segmentation.  The proposed method is able to learn to segment complex object
    classes using image annotation alone.
\item Providing a general and efficient open source software framework for
    structured prediction. 
\item Analysing max-margin learning algorithms with exact or approximate
    inference in different applications. We give a thorough evaluation of
    all major SSVM learning algorithms, using a wide array of application.
\item Showing that exact learning for semantic segmentation and object class
    segmentation is possible in practice, even in loopy graphs. We combine
    fast inference, caching and inference algorithms which certify optimality
    to learn a $1$-slack SSVM.
\end{itemize}

@@

We introduce an algorithm for semi-supervised learning of object class
segmentation in Chapter~\ref{ch:semi_supervised}, motivated by the problem of
annotating training data for semantic segmentation.

>>>>

%\includegraphics[width=\linewidth]{python_soul}%
%\flushright%
%Randall Munroe, \emph{xkcd.com}
\end{minipage}%
\\[2cm]
In computer vision research, the goal is to automatically extract information
from a given image or image sequence.  In particular discerning semantic
information, that is interpreting an image, is a prominent research topic.
While much progress has been made in recent years, computer vision systems still lag behind
human vision in most tasks that require semantic information. These tasks can often be formulated
in terms of \emph{semantic classes}, meaning categories of parts, objects or scenes.
Examples include answering questions such as ``Is this a picture of a beach?'', ``How many
cars are there in this image?'' or even ``What objects lie on the table?''.
These questions illustrate a range of possible tasks involving semantic categories,
such as classifying images of single objects, localization and counting of object classes,
and parsing a scene fully into objects and object classes together with their relations.
%
While humans can distinguish tens of thousands of object classes, and have
little trouble in interpreting complex scenes, current methods are often
restricted to a much smaller number of classes and only have limited
capabilities to model interactions or relations.  We believe that
\emph{context} is one of the most important cues when it comes to classifying
objects, and therefore understanding scenes. Therefore, we target dense
labeling of scenes, taking object relations into account.  The task of densely
labeling image pixels into classes is called object class segmentation or
semantic segmentation.
\pagebreak

We choose this task in particular for following reasons:
\begin{itemize*}
    \item Pixel-level annotation provides highly detailed information about the scene.
    \item Joint estimation of multiple classes allows for the use of context.
    \item In contrast to category-agnostic segmentation approaches, object class segmentation
        has an unambiguous true labeling.
    \item A variety of manually annotated datasets is publicly available.
\end{itemize*}
%
Applications of semantic segmentation and scene understanding include automatic
image interpretation for retrieval, autonomous driving and mobile robotics.
%While these are already important applications, we expect that with the
%abundance of camera data, and the proliferation of mobile computing, even
%more applications will emerge soon.
Besides these applications, due to the abundance of camera data and the
proliferation of mobile computing, we expect semantic annotation of images to
be a key component in future technologies.

@@

One of the bottlenecks in learning object class segmentation and semantic
segmentation is the availability of training data.  While unlabeled image data,
and even data with semantic
``tags'' is available in practically unlimited quantities, semantic annotation
on pixel level is scarce and only available through laborious manual annotation. 
Chapter~\ref{ch:semi_supervised} introduces an approach to cope with this
shortcoming of object class segmentation approaches by introducing a method to
learn segmentation automatically from image level annotations.

The main part of this thesis investigates the use of structured learning~\citep{taskar2003max, tsochantaridis2006large}
algorithms to the task of semantic image segmentation. Both topics have received
much attention in the computer vision and machine learning communities
lately~\citep{ladicky2009associative, krahenbuhl2012efficient,
branson2013efficient, blake2011markov}.
Unfortunately, learning structured prediction in computer vision applications
is still little understood.
We focus on the use of conditional random fields~(CRFs), which have shown
promising  results for computer vision applications. Using the paradigm of structural
support vector machines, it is possible to learn conditional random fields to
directly minimize a loss of interest. In particular, CRFs
allow to combine different cues, possibly produced using different paradgims in a
principled manner. One of the main difficulties with CRF approaches to computer
vision problems is that context in images is usually represented as a
neighborhood graph of pizels or superpixels.  These graphs, by nature, contain many cycles, making
inference intractable in general. Consequently, learning algorithms have to
rely on approximate inference, with often unknown consequences to learning.

@@

\section{List of Contributions}
This thesis contains the following contributions:
\begin{itemize}
\item Introducing a clustering algorithm that improves upon widely used
    approaches from the literature. Our algorithm yields better clusterings in
    terms of known classes on a wide range of standard datasets.
\item Demonstrating a scalable algorithm for weakly supervised object class
    segmentation.  The proposed method is able to learn to segment complex object
    classes using image annotation alone.
\item Providing a general and efficient open source software framework for
    structured prediction. 
\item Analysing max-margin learning algorithms with exact or approximate
    inference in different applications. We give a thorough evaluation of
    all major SSVM learning algorithms, using a wide array of application.
\item Showing that exact learning for semantic segmentation and object class
    segmentation is possible in practice, even in loopy graphs. We combine
    fast inference, caching and inference algorithms which certify optimality
    to learn a $1$-slack SSVM.
\item Learning 3D relations of semantic structure categories for indoor scenes.
    We extend the CRF approach to learning spacial relations from RGB-D data and
    improve upon the state-of-the-art in semantic annotation on the NYU V2 dataset
    of indoor scenes.
\end{itemize}

@@

We introduce an algorithm for semi-supervised learning of object class
segmentation in Chapter~\ref{ch:semi_supervised}, motivated by the difficulty of
obtraining annotated training data for semantic segmentation.

### 2013-09-23 itm/paper.tex
feedback from a real expert....

So not only is clustering one of the fundamental problems in machine learning, it is also
an important building block for the methods in this work.
%
Most algorithms are based on ad-hoc criteria such as intra-cluster similarity
and inter-cluster dissimilarity.
An alternative approach is to formalize clustering using an information
theoretic framework, where one considers inputs as well as cluster assignments
as random variables.  The goal is then to find an assignment of data points to
clusters that maximizes the mutual information between the assignments and the
observations.

In the following, we rely on a non-parametric estimator of the data entropy to find 
clusterings of maximum mutual information.
%
The use of non-parametric estimates allows a data-driven approach, without 
making strong assumptions on the form of the data distribution.
%
As a consequence, we obtain a very flexible model that, for example, allows 
non-convex clusters.
%
The resulting objective is easy to evaluate, but difficult to optimize
over. We overcome this by proposing an efficient approximate optimization
based on the Euclidean minimum spanning tree algorithm. 
%
Because the estimator and the optimization are both parameter-free, 
the only free parameter of the algorithm is the number of clusters, 
which makes it very easy to use in practice.
%
The contributions of this chapter are:
\begin{itemize}
    \item Proposing the use of a MST-based entropy estimator in information theoretic clustering.
    \item Giving a fast algorithm for a relaxed version of the resulting problem.
    \item Showing the practicality on a number of synthetic and real datasets.
    \item Extending the clustering to data on submanifolds by estimating implicit dimensionality.
\end{itemize}

>>>>

So not only is clustering one of the fundamental problems in machine learning, it is also
an important building block for other methods in this work.
%
Most algorithms are based on ad-hoc criteria such as intra-cluster similarity
and inter-cluster dissimilarity.
An alternative approach is to formalize clustering using an information
theoretic framework, where one considers inputs as well as cluster assignments
as random variables.  The goal is then to find an assignment of data points to
clusters that maximizes the mutual information between the assignments and the
observations.

In the following, we rely on a non-parametric estimator of the data entropy to find 
clusterings of maximum mutual information.
%
The use of non-parametric estimates allows a data-driven approach, without 
making strong assumptions on the form of the data distribution.
%
As a consequence, we obtain a very flexible model that, for example, allows 
non-convex clusters.
%
The resulting objective is easy to evaluate, but difficult to optimize
over. We overcome this by proposing an efficient approximate optimization
based on Euclidean minimum spanning trees.
%
Because the estimator and the optimization are both parameter-free, 
the only free parameter of the algorithm is the number of clusters, 
which makes it very easy to use in practice.
%
The non-parametric entropy estimate we
are using is applicable only if the data distribution is absolute continuous and
therefore can not be applied if the data lies on a proper submanifold.
We show how to overcome this obstacle in practice by using an estimate of the
intrinsic dimensionality of the data.
%
The contributions of this chapter are:
\begin{itemize}
    \item Proposing the use of a MST-based entropy estimator in information theoretic clustering.
    \item Giving a fast algorithm for a relaxed version of the resulting problem.
    \item Showing the practicality on a number of synthetic and real datasets.
    \item Extending the clustering to data on submanifolds by estimating intrinsic dimensionality.
\end{itemize}

### 2013-09-23 main.tex
fixes von riccarda, ein paar plots fuer auswertung
feedback from a real expert....

\begin{center}
\Large\sffamily
\vspace*{1cm}
\textbf{{\huge Methods for Learning Structured Prediction in Semantic Segmentation of Natural Images}}\\[5mm]
Dissertation\\
zur\\
Erlangung des Doktorgrades (Dr.\ rer.\ nat.)\\
der\\
Mathematisch-Naturwissenschaftlichen Fakult\"at\\
der\\
Rheinischen Friedrich-Wilhelm Universit\"at Bonn\\
vorgelegt von\\
{\LARGE Andreas Christian M\"uller}\\
aus\\
Offenbach am Main\\
Bonn, September 2013
\end{center}
\end{titlepage}
%
%\thispagestyle{empty}
\clearpage
\pagenumbering{Roman}
\setcounter{page}{2}
\vspace*{5cm}
\Large\sffamily
\begin{center}
Angefertigt mit Genehmigung\\
der Mathematisch-Naturwissenschaftlichen Fakult\"at\\
der Rheinischen Friedrich-Wilhelms-Universit\"at Bonn\\
\end{center}
1. Gutachter Prof. Dr. Sven Behnke\\
2. Gutachter Prof. Dr. Reinhard Klein\\
Tag der Promotion:\\
Erscheinungsjahr: 2013\\
\end{spacing}
\tableofcontents
%\listoffigures
%\listoftables
\chapter*{Zusammenfassung}
\enlargethispage{10mm}
\selectlanguage{german}
Automatische Segmentierung und Erkennung von semantischen Klassen in
nat\"urlichen Bildern ist ein wichtiges offenes Problem des maschinellen Sehens.
In dieser Arbeit untersuchen wir drei m\"oglichen Ans\"atze der Erkennung:
ohne \"Uberwachung, mit \"Uberwachung auf Ebene von Bildern und mit \"Uberwachung auf Ebene
von Pixeln.

@@

Der zweite Teil dieser Arbeit stellt ein Verfahren zur automatischen
Segmentierung und Erkennung von Objektklassen in nat\"urlichen Bildern vor, das
mit Hilfe von Supervision in Form von Klassen-Vorkommen auf Bildern in der Lage
ist ein Segmentierungsmodell zu lernen. Nach unserem Wissen ist dies der erste
Ansatz dieser Art, der sich speziell auf komplexe Objektklassen bezieht.

Der dritte und letzte Teil der Arbeit untersucht einen der am weitesten
verbreiteten Ans\"atze zur semantischen Segmentierung und
Objektklassensegmentierung, Conditional Random Fields, verbunden mit Verfahren
der strukturierten Vorhersage.
%Die Vorhersage f\"ur Modelle zur Bild-Markierung
%mittels Conditional Random Fields ist f\"ur gew\"ohnlich nur approiximativ
%moeglich, bedingt d\"urch das Vorhandensein von Kreisen in den zugrunde
%liegenden Nachbarschaftsgraphen.  Aus diesem Grund, und trotz der Beliebtheit
%von Conditional Random Field Verfahren, ist das strukturierte Lernen f\"ur
%solche Anwendungen schlecht verstanden.
%
Wir untersuchen verschiedene Lernalgorithmen des strukturierten Lernens, insbesondere
im Zusammenhang mit approximativer Inferenz. Wir zeigen, dass es m\"oglich ist trotz des Vorhandenseins von
Kreisen in den Nachbarschaftsgraphen exakte strukturierte Modelle zur Bildsegmentierung zu lernen.
Mit den vorgestellten Methoden bringen wir den Stand der Kunst
auf zwei komplexen Datens\"atzen zur semantischen Segmentierung voran, dem MSRC-21 Datensatz
von RGB-Bildern und dem NYU V2 Datensatz von RGB-D Bildern von Innenraum-Szenen.

>>>>

\begin{center}
\Large\sffamily
\vspace*{1cm}
\textbf{{\huge Methods for Learning Structured Prediction in Semantic Segmentation of Natural Images}}\\[5mm]
Dissertation\\
zur\\
Erlangung des Doktorgrades (Dr.\ rer.\ nat.)\\
der\\
Mathematisch-Naturwissenschaftlichen Fakult\"at\\
der\\
Rheinischen Friedrich-Wilhelm Universit\"at Bonn\\
vorgelegt von\\
{\LARGE Andreas Christian M\"uller}\\
aus\\
Offenbach am Main\\
Bonn, September 2013
\end{center}
\end{titlepage}
%
%\thispagestyle{empty}
\clearpage
\pagenumbering{Roman}
\setcounter{page}{2}
\vspace*{5cm}
\Large\sffamily
\begin{center}
Angefertigt mit Genehmigung\\
der Mathematisch-Naturwissenschaftlichen Fakult\"at\\
der Rheinischen Friedrich-Wilhelms-Universit\"at Bonn\\
\end{center}
1. Gutachter Prof. Dr. Sven Behnke\\
2. Gutachter Prof. Dr. Reinhard Klein\\
Tag der Promotion:\\
Erscheinungsjahr:\\
\end{spacing}
\tableofcontents
%\listoffigures
%\listoftables
\chapter*{Zusammenfassung}
\enlargethispage{10mm}
\selectlanguage{german}
Automatische Segmentierung und Erkennung von semantischen Klassen in
nat\"urlichen Bildern ist ein wichtiges offenes Problem des maschinellen Sehens.
In dieser Arbeit untersuchen wir drei m\"oglichen Ans\"atze der Erkennung:
ohne \"Uberwachung, mit \"Uberwachung auf Ebene von Bildern und mit \"Uberwachung auf Ebene
von Pixeln.

@@

Der zweite Teil dieser Arbeit stellt ein Verfahren zur automatischen
Segmentierung und Erkennung von Objektklassen in nat\"urlichen Bildern vor, das
mit Hilfe von Supervision in Form von Klassen-Vorkommen auf Bildern in der Lage
ist ein Segmentierungsmodell zu lernen.

Der dritte und letzte Teil der Arbeit untersucht einen der am weitesten
verbreiteten Ans\"atze zur semantischen Segmentierung und
Objektklassensegmentierung, Conditional Random Fields, verbunden mit Verfahren
der strukturierten Vorhersage.
%Die Vorhersage f\"ur Modelle zur Bild-Markierung
%mittels Conditional Random Fields ist f\"ur gew\"ohnlich nur approiximativ
%moeglich, bedingt d\"urch das Vorhandensein von Kreisen in den zugrunde
%liegenden Nachbarschaftsgraphen.  Aus diesem Grund, und trotz der Beliebtheit
%von Conditional Random Field Verfahren, ist das strukturierte Lernen f\"ur
%solche Anwendungen schlecht verstanden.
%
Wir untersuchen verschiedene Lernalgorithmen des strukturierten Lernens,
insbesondere im Zusammenhang mit approximativer Vorhersage. Wir zeigen, dass es
m\"oglich ist trotz des Vorhandenseins von Kreisen in den betrachteten Nachbarschaftsgraphen
exakte strukturierte Modelle zur Bildsegmentierung zu lernen.  Mit den
vorgestellten Methoden bringen wir den Stand der Kunst auf zwei komplexen
Datens\"atzen zur semantischen Segmentierung voran, dem MSRC-21 Datensatz von
RGB-Bildern und dem NYU V2 Datensatz von RGB-D Bildern von Innenraum-Szenen.

### 2013-09-23 nyu/paper.tex
feedback from a real expert....

%\begin{abstract}
%    We present a structured learning approach to semantic annotation of RGB-D
%    images. 
%    Our method learns to reason about spatial relations of objects and
%    fuses low-level class predictions to a consistent interpretation of a
%    scene. Our model incorporates color, depth and 3D scene features, on which
%    an energy function is learned to directly optimize object class prediction using
%    the loss-based maximum-margin principle of structural support vector machines.
%    %With the proliferation of cheap % TODO urghs formulation
%    %high-quality RGB-D sensors, it became possible to capture more and more
%    %data not only as color images, but with additional depth information. The
%    %depth channel provides important clues not only for scene structure, but
%    %also for predicting semantic object classes.
%    %We demonstrate how CRF learning, an approach
%    %that has shown great promise for RGB images, can exploit depth information
%    %to improve semantic interpretation of the scene.
%    We evaluate our approach on the NYU V2 dataset of indoor scenes, a challenging
%    dataset covering a wide variety of scene layouts and object classes. We
%    hard-code much less information about the scene layout into our model then previous
%    approaches, and instead learn object relations directly from the data. We
%    find that our conditional random field approach improves upon previous
%    work, setting a new state-of-the-art for the dataset.
%\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%\section{Introduction}
For robots to perform varied tasks in unstructured environments, understanding their
surroundings is essential. We formulate the problem of semantic annotation of maps
as a dense labeling of RGB-D images into semantic classes. Dense labeling of objects
and structure classes allows for a detailed reasoning about the scene.

In this work, we propose the use of random forests combined with conditional
random fields (CRFs) to perform robust estimation of structure classes in RGB-D
images. The CRF is learned using a structural support vector machine,
allowing it to integrate the noisy low-level information produced by a pixel-based
random forest to a consistent interpretation of the scene.

We thereby extend the success of learned CRF models for
semantic segmentation in RGB images to the domain of 3D scenes.
Our emphasis lies on exploiting the additional depth and 3D information in all
processing steps, while relying on \emph{learning} to create a model that 
is adjusted to the properties of the sensor input and environment.

Our approach starts with a random forest, providing a noisy local estimate
of semantic classes based on color and depth information. These estimates
are grouped together using a superpixel approach, for which we extend previous
superpixel algorithms from the RGB to the RGB-D domain.
We then build a geometric model of the scene, based on the neighborhood graph
of superpixels.  We use this graph not only to capture spatial relations in the
2D plane of the image, but also to model object distances and surface angles in
3D, using a point cloud generated from the RGB-D image. The process is depicted
in \Figref{teaser}.

@@

Recent approaches for indoor semantic annotation of maps mostly focused on
RGB-D images, which are now easy to obtain using structured light sensors.
\citet{stuckler2012semantic}, for example, used a Random Forest model to obtain a dense
semantic labeling of images and integrated predictions over multiple views in 3D.
They evaluated their approach on table-top and simple indoors scenes.
%
\citet{silberman2011indoor} introduced the NYU Depth Dataset V1 indoor dataset,
which consisted of a large variety of densely annotated indoor scenes, captured as RGB-D images.
Their work also introduced a baseline method for semantic segmentation of RGB-D
image, which is based on a CRF over superpixels, with unary potentials given by interest point descriptors.
While pairwise potentials for the CRF were carefully
designed for the dataset, potentials were either directly set by hand or estimated using empirical
frequencies. This is in contrast to our work, which applies structured prediction techniques
to learn potentials automatically to optimize predictive performance.
%
\citet{ren2012rgb} evaluated the design of features for semantic labeling of
RGB-D data, and use a hierarchical segmentation to provide context. While they
also define a CRF on superpixels, their model is again not learned, but a
weighted Potts model, using only a probability of boundary map, and not taking
spatial layout into account at all.
%
\citet{SilbermanECCV12} extended the NYU Depth Dataset V1 to the NYU Depth Dataset V2 that we are also
using in this work.  Their focus is on inferring support relations in indoor
scenes, such as objects resting on tables or shelfs, which in turn rest on the floor.
Their approach is based on robust estimation of 3D plane hypotheses, which are then jointly
optimized with support relations and structure classes.
\citet{SilbermanECCV12} used a complex pipeline, employing significant domain knowledge.
In this work, on the other hand, we try to learn all relevant domain specific features
directly from the data, which allows us to out-perform the work of \citet{SilbermanECCV12}
with respect to structure class segmentation.

\citet{couprie-iclr-13} approached the task of semantic segmentation of structure classes
in RGB-D using the paradigm of convolutional neural networks, extending
previous work of \citet{farabet-pami-13} and \citet{schulz2012learning}.
Similar to our approach, \citet{couprie-iclr-13} combined the output of a
pixel-based, low-level learning algorithm with an independent unsupervised
segmentation step. In contrast to their work, we improve our results by not
only averaging predictions within superpixels, but also explicitly learning
interactions between neighboring superpixels, favoring a consistent
interpretation of the whole image.

@@

\subsection{Unary Image Features}\seclabel{unaries}
Our method builds on the probability output of a random forest, trained for pixel-wise
classification of the structure classes.
We use the GPU implementation provided by \citet{stueckler2013}\footnote{\url{https://github.com/deeplearningais/curfil}}.
The input for training are the full RGB-D images, transformed to Lab
color space. Each tree in the forest uses training pixels only from a subset of
training images. For each training image, an equal number of pixels for each
occurring class is sampled. Split features are given by difference of regions
on color or depth channels. Region size and offsets are normalized using the
depth at the target pixel.
We accumulate the probabilistic output for all pixels within a superpixel, and
use the resulting distribution as a feature for the unary node potentials in our CRF model.
We augment these prediction with another feature, based on the height of a superpixel in 3D.
This is a very informative feature, in particular to determine the floor.
To compute the height of a (super) pixel, we first find the ``up'' direction.
We use a very simple approach that we found effective: we cluster normal
directions of all pixels into 10 clusters using $k$-means, and use the one that
is most parallel to the $Y$ direction, which roughly corresponds to height in
the dataset.
We then project the 3D point cloud given by the depth along this direction, and
normalize the result between $0$ and $1$.
This procedure works robustly given there is some horizontal surface in the image,
such as the ground or a table. A few scenes contain only walls and furniture,
and the approach fails for these. \Figref{height} illustrates a typical case and
one of the much rarer failure cases.
While we could use a more elaborate scheme,
such as the one from \citet{SilbermanECCV12}, we suspect that the feature is of
little use in scenes without horizontal surfaces.
\begin{figure}
    \begin{center}
        \includegraphics[width=\linewidth]{nyu/images/normal_feature}
    \end{center}
    \caption{%
        Visualization of one of the pairwise features, the similarity between superpixel normals.
        The image shows the zoom-in of a bedroom scene, together with the superpixel over-segmentation.
        Lines connect adjacent superpixels, and line-strength gives the magnitude of the orientation similarity.
    \figlabel{normal_feature}}
\end{figure}

@@

We evaluate our approach on the public NYU depth V2 segmentation dataset of
indoor scenes.  The dataset comes with a detailed annotation of 1449 indoor
RGB-D images belonging to a wide variety of indoor scenes, categorized into 26
scene classes.  The annotation contains four semantic structural classes:
structure, floor, furniture and prop. There is an additional ``void'' class,
which is used for object boundaries and hard-to-annotate regions. We follow the
literature in excluding these pixels completely from the evaluation.
We optimize our model for \emph{average class accuracy} (the mean of the
diagonal of the confusion matrix), which puts more emphasis on the harder
classes of props and furniture, which have smaller area than structure and
floor.

@@

\section{Summary and Discussion}
We introduce a CRF formulation for semantic segmentation
of structure classes in RGB-D images. We base our model in the output of an efficient GPU implementation
of random forest, and model spatial neighborhood using a superpixel-based approach.
We combine color, depth and 3D orientation features into an energy function
that is learned using the SSVM approach. By explicitly modeling 3D relations in
a fully learned framework, we improve the state-of-the art on the NYU V2
dataset for semantic annotation of structure classes.

While our approach allows modeling of spatial relations, these are limited to
local interactions. In future work, these interactions could be extended to
larger areas using latent variable models or higher
order potentials~\citep{ladicky2009associative}.
Another possible line of future work is to combine our approach with a more
task-specific one, directly including support plane assumptions into the model,
as done by \citet{SilbermanECCV12}. Finally, we could also combine our
single-frame approach with the approach of \citet{stueckler2013}, which fuses
individual views in 3D to exploit temporal coherence.

While this work did not explicitly address real time application, the random
forest implementation that we build upon allows for real-time
processing~\citep{stueckler2013}. The SLIC superpixel algorithm can also be
implemented on GPU in real-time, as was demonstrated by \citep{ren2011gslic}, and similarly
the normal features we use also have
real-time capabilities~\citep{holz_2011_robocup}.
Finally, fusion move inference for our model is very efficient for our model,
opening up the possibility to implement our approach entirely in real time.

>>>>

@@

%\section{Introduction}
For robots to perform varied tasks in unstructured environments, understanding their
surroundings is essential. In this chapter, we look at the semantic annotation of maps
as a dense labeling of \mbox{RGB-D} images into semantic classes. We formulate the problem
as learning a CRF over a superpixel segmentation of the \mbox{RGB-D} image, producing a labeling
that takes 3D layout into account. Dense labeling of objects
and structure classes allows for a detailed reasoning about the scene.

%We use of random forests to obtain a noisy low-level estimate of structure class on pixel level.
%These estimates serve as the unary input to the learned CRF model which integrates
%these low-level cues to a consistent interpretation of the scene.

We thereby extend the success of learned CRF models for semantic segmentation
in RGB images as considered in Chapter~\ref{exact_learning} to the domain of 3D scenes.
Our emphasis lies on exploiting the additional depth and 3D information in all
processing steps, while relying on \emph{learning} to create a model that 
is adjusted to the properties of the sensor input and environment.

Our approach starts with a random forest, providing a noisy local estimate
of semantic classes based on color and depth information. These estimates
are grouped together using a superpixel approach, for which we extend previous
superpixel algorithms from the RGB to the \mbox{RGB-D} domain.
We then build a geometric model of the scene, based on the neighborhood graph
of superpixels.  We use this graph not only to capture spatial relations in the
2D plane of the image, but also to model object distances and surface angles in
3D, using a point cloud generated from the \mbox{RGB-D} image. The process is depicted
in \Figref{teaser}.

@@

Recent approaches for indoor semantic annotation of maps mostly focused on
\mbox{RGB-D} images, which are now easy to obtain using structured light sensors.
\citet{stuckler2012semantic}, for example, used a Random Forest model to obtain a dense
semantic labeling of images and integrated predictions over multiple views in 3D.
They evaluated their approach on table-top and simple indoors scenes.
%
\citet{silberman2011indoor} introduced the NYU Depth Dataset V1 indoor dataset,
which consisted of a large variety of densely annotated indoor scenes, captured as \mbox{RGB-D} images.
Their work also introduced a baseline method for semantic segmentation of \mbox{RGB-D}
image, which is based on a CRF over superpixels, with unary potentials given by interest point descriptors.
While pairwise potentials for the CRF were carefully
designed for the dataset, potentials were either directly set by hand or estimated using empirical
frequencies. This is in contrast to our approach which applies structured prediction techniques
to learn potentials automatically to optimize predictive performance.
%
\citet{ren2012rgb} evaluated the design of features for semantic labeling of
\mbox{RGB-D} data, and use a hierarchical segmentation to provide context. While they
also define a CRF on superpixels, their model is again not learned, but a
weighted Potts model, using only a probability of boundary map, and not taking
spatial layout into account at all.
%
\citet{SilbermanECCV12} extended the NYU Depth Dataset V1 to the NYU Depth Dataset V2 that we are also
using here.  Their focus is on inferring support relations in indoor
scenes, such as objects resting on tables or shelfs, which in turn rest on the floor.
Their approach is based on robust estimation of 3D plane hypotheses, which are then jointly
optimized with support relations and structure classes.
\citet{SilbermanECCV12} used a complex pipeline, employing significant domain knowledge.
In our approach, on the other hand, we try to learn all relevant domain specific features
directly from the data, which allows us to out-perform the work of \citet{SilbermanECCV12}
with respect to structure class segmentation.

\citet{couprie-iclr-13} approached the task of semantic segmentation of structure classes
in \mbox{RGB-D} using the paradigm of convolutional neural networks, extending
previous work of \citet{farabet-pami-13} and \citet{schulz2012learning}.
Similar to our approach, \citet{couprie-iclr-13} combined the output of a
pixel-based, low-level learning algorithm with an independent unsupervised
segmentation step. In contrast to their work, we improve our results by not
only averaging predictions within superpixels, but also explicitly learning
interactions between neighboring superpixels, favoring a consistent
interpretation of the whole image.

@@

\subsection{Unary Image Features}\seclabel{unaries}
Our method builds on the probability output of a random forest, trained for pixel-wise
classification of the structure classes.
We use the GPU implementation provided by \citet{stueckler2013}\footnote{\url{https://github.com/deeplearningais/curfil}}.
The input for training are the full \mbox{RGB-D} images, transformed to Lab
color space. Each tree in the forest uses training pixels only from a subset of
training images. For each training image, an equal number of pixels for each
occurring class is sampled. Split features are given by difference of regions
on color or depth channels. Region size and offsets are normalized using the
depth at the target pixel.
We accumulate the probabilistic output for all pixels within a superpixel, and
use the resulting distribution as a feature for the unary node potentials in our CRF model.
We augment these prediction with another feature, based on the height of a superpixel in 3D.
This is a very informative feature, in particular to determine the floor.
To compute the height of a (super) pixel, we first find the ``up'' direction.
We use a very simple approach that we found effective: we cluster normal
directions of all pixels into 10 clusters using $k$-means, and use the one that
is most parallel to the $Y$ direction, which roughly corresponds to height in
the dataset.
We then project the 3D point cloud given by the depth along this direction, and
normalize the result between $0$ and $1$.
This procedure works robustly given there is some horizontal surface in the image,
such as the ground or a table. A few scenes contain only walls and furniture,
and the approach fails for these. \Figref{height} illustrates a typical case and
one of the much rarer failure cases.
While we could use a more elaborate scheme,
such as the one from \citet{SilbermanECCV12}, we suspect that the feature is of
little use in scenes without horizontal surfaces.
\begin{figure}
    \begin{center}
        \includegraphics[width=\linewidth]{nyu/images/normal_feature}
    \end{center}
    \caption{%
        Visualization of one of the pairwise features, the similarity between superpixel normals.
        The image shows the zoom-in of a bedroom scene, together with the superpixel over-segmentation.
        Lines connect adjacent superpixels, and line-strength gives the magnitude of the orientation similarity.
    \figlabel{normal_feature}}
\end{figure}

@@

We evaluate our approach on the public NYU depth V2 segmentation dataset of
indoor scenes.  The dataset comes with a detailed annotation of 1449 indoor
\mbox{RGB-D} images belonging to a wide variety of indoor scenes, categorized into 26
scene classes.  The annotation contains four semantic structural classes:
structure, floor, furniture and prop. There is an additional ``void'' class,
which is used for object boundaries and hard-to-annotate regions. We follow the
literature in excluding these pixels completely from the evaluation.
We optimize our model for \emph{average class accuracy} (the mean of the
diagonal of the confusion matrix), which puts more emphasis on the harder
classes of props and furniture, which have smaller area than structure and
floor.

@@

\section{Summary and Discussion}
We introduce a CRF formulation for semantic segmentation
of structure classes in \mbox{RGB-D} images. We base our model in the output of an efficient GPU implementation
of random forest, and model spatial neighborhood using a superpixel-based approach.
We combine color, depth and 3D orientation features into an energy function
that is learned using the SSVM approach. By explicitly modeling 3D relations in
a fully learned framework, we improve the state-of-the art on the NYU V2
dataset for semantic annotation of structure classes.

While our approach allows modeling of spatial relations, these are limited to
local interactions. In future work, these interactions could be extended to
larger areas using latent variable models or higher
order potentials~\citep{ladicky2009associative}.
Another possible line of future investigation is to combine our approach with a more
task-specific one, directly including support plane assumptions into the model,
as done by \citet{SilbermanECCV12}. Finally, we could also combine our
single-frame approach with the approach of \citet{stueckler2013}, which fuses
individual views in 3D to exploit temporal coherence.

We did not explicitly address real time application, the random
forest implementation that we build upon allows for real-time
processing~\citep{stueckler2013}. The SLIC superpixel algorithm can also be
implemented on GPU in real-time, as was demonstrated by \citep{ren2011gslic}, and similarly
the normal features we use also have
real-time capabilities~\citep{holz_2011_robocup}.
Finally, fusion move inference for our model is very efficient for our model,
opening up the possibility to implement our approach entirely in real time.
